Title Goes Here

### 2017-08-17 analysis/paper/paper.Rmd
writing about the JAS survey

# Email survey

@@

We conducted a pilot study to explore data sharing in archaeology. Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`. We focused on stone artefact experiments because the all research materials are generated by the researcher so there are not other stakeholders to consider when determining how to share the data: the researcher has complete control over the data. There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects. In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress). We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.

We sent out `r emails_out` emails to authors of these papers, and received `r emails_back`, resulting in a `r email_response_rate`% response rate. We recieved `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%. Wi

change over time...

location...
student or not...

# Background

>>>>

# Email survey of data sharing behaviours

@@

We conducted a small pilot study to explore data sharing in archaeology. Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`. We focused on stone artefact experiments because the all research materials are generated by the researcher so there are not other stakeholders to consider when determining how to share the data: the researcher has complete control over the data. There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects. In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress). We promised to contact the authors if we discovered anything new, and not to share the data further without their permission. The data included with this paper is anonymised so the specific authors we contacted cannot be identified.

We sent out `r emails_out` emails to authors of these papers, and received `r emails_back`, resulting in a `r email_response_rate`% response rate. We recieved `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%. Our small sample size limits the robustness of our analysis, but points to some observations that may be worth following up in a larger study. We found no clear relationship of data sharing and date of publication. We do not have any insights into why authors of older papers decline to share, but three of the six responses for papers published in 2014-2015 said they declined to share data because they intend to use these data in a future publication. Other noteworthy reasons provided by respondents for delining to share include the author about to defend their thesis or about to go on vacation. We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised for ready for distribution. We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing. Similarly there are no clear patterns data sharing of among different journals. It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research."

```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication", eval = TRUE}
library(viridis)
years <- seq(min(email_survey$'publication year', na.rm = TRUE),
    max(email_survey$'publication year', na.rm = TRUE), 1)
email_survey %>% 
  filter(!is.na( `email sent date`)) %>% 
  group_by(`publication year`, `content recieved date`) %>% 
  tally() %>% 
  filter(!is.na(`publication year`)) %>% 
  mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
  "yes", "no")) %>% 
  ggplot(aes(`publication year`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
    scale_fill_viridis(discrete = TRUE,
                       guide = guide_legend(title = "Data shared?")) +
    scale_x_continuous(breaks = years,
                       labels = years) +
    theme_bw()

```{r student, fig.cap="No relationship between student status and data sharing behaviour", eval = TRUE}
email_survey %>% 
  filter(!is.na( `email sent date`)) %>% 
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
  "yes", "no")) %>% 
  group_by(
  `author is graduate student at time of publication?`, 
  `content recieved date`) %>% 
  tally() %>% 
  filter(!is.na(`author is graduate student at time of publication?`)) %>% 
  ggplot(aes(`author is graduate student at time of publication?`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
    scale_fill_viridis(discrete = TRUE,
                       guide = guide_legend(title = "Data shared?")) +
    theme_bw()
```

```{r author-country, eval = TRUE}
email_survey %>% 
  filter(!is.na( `email sent date`)) %>% 
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
  "yes", "no")) %>% 
  group_by(
  `author home`, 
  `content recieved date`) %>% 
  tally() %>% 
  filter(!is.na(`author home`)) %>% 
  ggplot(aes(`author home`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
    scale_fill_viridis(discrete = TRUE,
                       guide = guide_legend(title = "Data shared?")) +
    theme_bw()
```

```{r journals, fig.cap="No relationship between sharing behaviour and journal.", eval = TRUE}
email_survey %>% 
  filter(!is.na( `email sent date`)) %>% 
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
  "yes", "no")) %>% 
  group_by(
  journal, 
  `content recieved date`) %>% 
  tally() %>% 
  filter(!is.na(journal)) %>% 
  ggplot(aes(journal, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
    scale_fill_viridis(discrete = TRUE,
                       guide = guide_legend(title = "Data shared?")) +
    theme_bw()
```

# Survey of articles published in JAS

```{r eval = TRUE}
jas_article_survey <- readr::read_csv("../data/raw_data/jas_article_survey.csv")

have_data_avail_statement <- nrow(jas_article_survey[jas_article_survey$`Data availability statement` != "none", ])
have_data_avail_statement_perc <- round(have_data_avail_statement / nrow(jas_article_survey) * 100,0)
data_are_avail <- nrow(jas_article_survey[jas_article_survey$`Can I access the data without contacting the authors?` %in% c("Yes", "yes"), ])
data_are_avail_perc <- round(data_are_avail / nrow(jas_article_survey) * 100,0)

data_in_table <-  nrow(jas_article_survey[grepl("Tab", jas_article_survey$`Where is the data?`), ])
data_in_table_perc <- round(data_in_table / data_are_avail * 100, 0)

data_in_supp <-  nrow(jas_article_survey[grepl("Sup", jas_article_survey$`Where is the data?`), ])
data_in_supp_perc <- round(data_in_supp / data_are_avail * 100, 0)

data_in_repo <-  nrow(jas_article_survey[grepl("www|dryad|osf", tolower(jas_article_survey$`Where is the data?`)), ])
data_in_repo_perc <- round(data_in_repo / data_are_avail * 100, 0)

data_type <- jas_article_survey %>% 
  filter(`data type?` != "na") %>% 
  group_by(`data type?`) %>% 
  tally(sort = TRUE) %>% 
  mutate(perc = round(n / sum(n) *100,1))

comp_n <- data_type %>% 
  filter(`data type?`  == "compositional") %>% 
  pull(n)

comp_perc <- data_type %>% 
  filter(`data type?`  == "compositional") %>% 
  pull(perc)

dna_n <- data_type %>% 
  filter(`data type?`  == "DNA sequence") %>% 
  pull(n)

dna_perc <- data_type %>% 
  filter(`data type?`  == "DNA sequence") %>% 
  pull(perc)

dates_n <- data_type %>% 
  filter(grepl("dates", `data type?`)) %>% 
  summarise(n = sum(n)) %>% 
  pull(n)

dates_perc <- data_type %>% 
  filter(grepl("dates", `data type?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(perc)

file_type <- jas_article_survey %>% 
  filter(`What file format?` != "na") %>% 
  group_by(`What file format?`) %>% 
  tally(sort = TRUE) %>% 
  mutate(perc = round(n / sum(n) *100,1))

xlsx_n <- file_type %>% 
  filter(grepl("XLSX", `What file format?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(n)

xlsx_perc <- file_type %>% 
  filter(grepl("XLSX", `What file format?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(perc)

csv_n <- file_type %>% 
  filter(grepl("CSV", `What file format?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(n)

csv_perc <- file_type %>% 
  filter(grepl("CSV", `What file format?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(perc)

doc_n <- file_type %>% 
  filter(grepl("doc|DOC", `What file format?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(n)

doc_perc <- file_type %>% 
  filter(grepl("doc|DOC", `What file format?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(perc)

software <- jas_article_survey %>% 
  filter(`Analysis software?` != "na") %>% 
  group_by(`Analysis software?`) %>% 
  tally(sort = TRUE) %>% 
  mutate(perc = round(n / sum(n) *100,1))

script_n <- software %>% 
  filter(grepl("^R$|.*R$|Python|MATLAB|OxCal", `Analysis software?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(n)

script_perc <- software %>% 
  filter(grepl("^R$|.*R$|Python|MATLAB|OxCal", `Analysis software?`)) %>% 
  summarise(n = sum(n),
            perc = sum(perc)) %>% 
  pull(perc)

In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` articles published during 2016-2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviours. In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement. For papers where data was available, `r data_in_supp` papers (`r data_in_supp_perc`%) have data contained in supplementary files published with the paper, `r data_in_table` (`r data_in_table_perc`%) have all the raw data in tables in the text of the paper, and `r data_in_repo` (`r data_in_repo_perc`%) have data in an online repository. The most frequently shared type of data is compositional (e.g. element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%). Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, a open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%). When a scripting language such as R, Python, MATLAB or OxCal was used as the analysis software for a paper (n = `r script_n`, `r script_perc`% of the papers where would identify the software), only 2 papers also provided the script with the paper.

correlation of data type and location of sharing

# Data sets in repositories
